Understanding the Lomb-Scargle Periodogram

### 2017-02-03 PracticalLombScargle.tex
Add Koen 2006 reference

We should keep in mind one consequence of this Nyquist definition:
if you have any pair of observation spacings
whose ratio is irrational, {\it the Nyquist limit does not exist!}.
To realize this situation in practice, however, 
would require infinitely precise measurements of the
times $t_i$; finite precision of time measurements means the Nyquist
frequency can be large, but not infinite.
For example, if your observation times are recorded to $D$ decimal places,
the Nyquist frequency will be at most
\begin{equation}
  f_{Ny} \le \frac{1}{2} 10^D,
  \eqlabel{nonuniform-nyquist}
\end{equation}
with the inequality due to the fact that larger common factors may exist.
In other words, absent other relevant patterns in the observations,
the Nyquist frequency for irregularly-sampled data is most
typically set by {\it the precision of the time measurements}.

@@

One important caveat of the Lomb-Scargle approach is that the statistical
guarantees only apply when the observations have {\it uncorrelated white
noise}; data with more complicated noise characteristics must be treated
more carefully; see, e.g., \citep{Vio2010} or the Least Squares approach
discussed in \sect{extensions-observational-noise}.

@@

This also suggests a convenient way to construct a periodogram that accounts
for correlated noise across observations; denoting the  $N\times N$ noise
covariance matrix for $N$ observations by $\Sigma$, and constructing the
vectors $\vec{y} = [y_1, y_2,\cdots y_n]^T$
and $\vec{y}_{model} = [y_{model}(t_1),y_{model}(t_1),\cdots y_{model}(t_n)]^T$,
then the $\chi^2$ statistic can be written
\begin{equation}
  \chi^2(f) = (\vec{y}-\vec{y}_{model})^T\Sigma^{-1}(\vec{y}-\vec{y}_{model})
  \eqlabel{chi2-with-correlated-errors}
\end{equation}
In the case of uncorrelated errors, the off-diagonal terms of $\Sigma$
are zero, and this expression reduces to \eq{chi2-with-errors}.
This is similar in spirit to the approach to correlated noise developed
by \citet{Vio2010} from the Fourier perspective, and is in fact identical
in the case of the ``zero-mean colored noise'' example considered therein.

>>>>

We should keep in mind one consequence of this Nyquist definition:
if you have any pair of observation spacings
whose ratio is irrational, {\it the Nyquist limit does not exist!}.
To realize this situation in practice, however, 
would require infinitely precise measurements of the
times $t_i$; finite precision of time measurements means the Nyquist
frequency can be large, but not infinite.
For example, if your observation times are recorded to $D$ decimal places,
the Nyquist frequency will be at most
\begin{equation}
  f_{Ny} \le \frac{1}{2} 10^D,
  \eqlabel{nonuniform-nyquist}
\end{equation}
with the inequality due to the fact that larger common factors may exist.
In other words, absent other relevant patterns in the observations,
the Nyquist frequency for irregularly-sampled data is most
typically set by {\it the precision of the time measurements}
\citep[see][for a more rigorous derivation of this]{Koen2006}.

@@

One important caveat of the simple Lomb-Scargle formula is that the statistical
guarantees only apply when the observations have {\it uncorrelated white
noise}; data with more complicated noise characteristics must be treated
more carefully; see, e.g., \citet{Vio2010} or the Least Squares approach
discussed in \sect{extensions-observational-noise}.

@@

This generalization of $\chi^2(f)$ also suggests a convenient way to
construct a periodogram in the presence of correlated observational noise.
If we let $\Sigma$ denote the  $N\times N$ noise
covariance matrix for $N$ observations, and construct the
vectors
\begin{eqnarray}
  &\vec{y} = [y_1, y_2,\cdots y_n]^T \nonumber\\
  &\vec{y}_{model} = [y_{model}(t_1),y_{model}(t_1),\cdots y_{model}(t_n)]^T,
\end{eqnarray}
then the $\chi^2$ expression for correlated errors can be written
\begin{equation}
  \chi^2(f) = (\vec{y}-\vec{y}_{model})^T\Sigma^{-1}(\vec{y}-\vec{y}_{model}),
  \eqlabel{chi2-with-correlated-errors}
\end{equation}
which reduces to \eq{chi2-with-errors} if noise is uncorrelated (i.e., if the
off-diagonal terms of $\Sigma$ are zero).
This resulting periodogram is quite similar to the approach to correlated
noise developed by \citet{Vio2010} from the Fourier perspective,
and is in fact exactly equivalent in the case of the
``zero-mean colored noise'' example considered therein.
